---
title: "FUT2daa"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DAA
Perform differential abundance analysis on the FUT2 microbiome data.
The analysis will be done on the first two timepoints only (Early and Mid) as by the third timepoint many of the patients have delivered).

### Analyses to perform:
1. Comparison of the microbial composition between secretors and non-secretors at timepoint 1.
2. Comparison of the microbial composition between secretors and non-secretors at timepoint 2.
3. Comparison of the microbial composition between the LD- secretors and the LD- non-secretors (timepoint 1).
3. Comparison of the microbial composition between the LD- secretors and the LD- non-secretors (aggregate model - take the lowest LD sample for each pregnancy [between timepoint 1 and 2]).

```{r}
library(ggplot2)
library(ALDEx2)
#import microbial species counts and assign secretor status to the samples
t1asv <- read.csv("T1CountsDAA.csv", header=T, row.names=1)
t1conds <- c(rep("NSec",77), rep("Sec",210))
t2asv <- read.csv("T2CountsDAA.csv", header=T, row.names=1)
t2conds <- c(rep("NSec",75), rep("Sec",206))
t1.aldex <- aldex.clr(t1asv, t1conds, mc.samples=1000, verbose=T)
t1.aldex.tt <- aldex.ttest(t1.aldex, paired.test=F, verbose=T)
t1.aldex.effect <- aldex.effect(t1.aldex, verbose=F)
t1.all <- data.frame(t1.aldex.tt, t1.aldex.effect, stringsAsFactors = F)
t2.aldex <- aldex.clr(t2asv, t2conds, mc.samples=1000, verbose=T)
t2.aldex.tt <- aldex.ttest(t2.aldex, paired.test=F, verbose=T)
t2.aldex.effect <- aldex.effect(t2.aldex, verbose=F)
t2.all <- data.frame(t2.aldex.tt, t2.aldex.effect, stringsAsFactors = F)
```

### Comparison of secretors and non-secretors at Early pregnancy
No individual organisms are significantly differentially abundant and the effect sizes are very small.

Plot as horizontal barplot in order of effect size

```{r}
par(mfrow=c(1,2))
aldex.plot(t1.all, type="MA", test="welch")
aldex.plot(t1.all, type="MW", test="welch")
t1.all$spp <- with(t1.all, reorder(rownames(t1.all), effect))
#t1.all.order <- t1.all[order(t1.all$effect),]
t1.all[["sign"]] = ifelse(t1.all[["effect"]] >= 0, "secretor", "non-secretor")
par(mfrow=c(1,1))
t1.effect.plot <- ggplot(t1.all, aes(x=spp, y=effect, fill=sign))+geom_bar(stat="identity", width=0.5)+coord_flip()+theme_classic()+scale_fill_discrete(name="Enriched")
t1.effect.plot
```

### Comparison of secretors and non-secretors at Mid pregnancy
The timepoint 2 pregnancies show a similar profile with respect to an increase in the amount of L.iners in the non-secretors.

```{r}
par(mfrow=c(1,2))
aldex.plot(t2.all, type="MA", test="welch")
aldex.plot(t2.all, type="MW", test="welch")
t2.all$spp <- with(t2.all, reorder(rownames(t2.all), effect))
t2.all[["sign"]] = ifelse(t2.all[["effect"]] >= 0, "secretor", "non-secretor")
par(mfrow=c(1,1))
t2.effect.plot <- ggplot(t2.all, aes(x=spp, y=effect, fill=sign))+geom_bar(stat="identity", width=0.5)+coord_flip()+theme_classic()+scale_fill_discrete(name="Enriched")
t2.effect.plot
```

### DAA plots in early pregnancy 

```{r}
par(mfrow=c(1,2))
aldex.plot(t1DEP.all, type="MA", test="welch")
aldex.plot(t1DEP.all, type="MW", test="welch")
t1DEP.all$spp <- with(t1DEP.all, reorder(rownames(t1DEP.all), effect))
t1DEP.all[["sign"]] = ifelse(t1DEP.all[["effect"]] >= 0, "secretor", "non-secretor")
par(mfrow=c(1,1))
t1DEP.effect.plot <- ggplot(t1DEP.all, aes(x=spp, y=effect, fill=sign))+geom_bar(stat="identity", width=0.5)+coord_flip()+theme_classic()+scale_fill_discrete(name="Enriched")
t1DEP.effect.plot
```

### DAA Plots in mid pregnancy

```{r}
par(mfrow=c(1,2))
aldex.plot(t2DEP.all, type="MA", test="welch")
aldex.plot(t2DEP.all, type="MW", test="welch")
t2DEP.all$spp <- with(t2DEP.all, reorder(rownames(t2DEP.all), effect))
t2DEP.all[["sign"]] = ifelse(t2DEP.all[["effect"]] >= 0, "secretor", "non-secretor")
par(mfrow=c(1,1))
t2DEP.effect.plot <- ggplot(t2DEP.all, aes(x=spp, y=effect, fill=sign))+geom_bar(stat="identity", width=0.5)+coord_flip()+theme_classic()+scale_fill_discrete(name="Enriched")
t2DEP.effect.plot
```

