---
title: "FUT2daa"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DAA
Perform differential abundance analysis on the FUT2 microbiome data

### Analyses to perform:
1. Comparison of the microbial composition between secretors and non-secretors at timepoint 1.
2. Comparison of the microbial composition between secretors and non-secretors at timepoint 2.
3. Comparison of the microbial composition between the LD- secretors and the LD- non-secretors (timepoint 1).
3. Comparison of the microbial composition between the LD- secretors and the LD- non-secretors (aggregate model - take the lowest LD sample for each pregnancy [between timepoint 1 and 2]).

```{r}
library(ggplot2)
library(ALDEx2)
t1asv <- read.csv("T1CountsDAA.csv", header=T, row.names=1)
t1conds <- c(rep("NSec",77), rep("Sec",210))
t2asv <- read.csv("T2CountsDAA.csv", header=T, row.names=1)
t2conds <- c(rep("NSec",75), rep("Sec",206))
t1.aldex <- aldex.clr(t1asv, t1conds, mc.samples=1000, verbose=T)
t1.aldex.tt <- aldex.ttest(t1.aldex, paired.test=F, verbose=T)
t1.aldex.effect <- aldex.effect(t1.aldex, verbose=F)
t1.all <- data.frame(t1.aldex.tt, t1.aldex.effect, stringsAsFactors = F)
t2.aldex <- aldex.clr(t2asv, t2conds, mc.samples=1000, verbose=T)
t2.aldex.tt <- aldex.ttest(t2.aldex, paired.test=F, verbose=T)
t2.aldex.effect <- aldex.effect(t2.aldex, verbose=F)
t2.all <- data.frame(t2.aldex.tt, t2.aldex.effect, stringsAsFactors = F)
```

### Comparison of secretors and non-secretors at Timepoint 1
No individual organisms are significantly differentially abundant though the effect sizes indicate that the most differences occur with respect to Lactobacillus iners (slightly enriched in the non-secretors).

Plot as horizontal barplot in order of effect size

```{r}
par(mfrow=c(1,2))
aldex.plot(t1.all, type="MA", test="welch")
aldex.plot(t1.all, type="MW", test="welch")
t1.all$spp <- with(t1.all, reorder(rownames(t1.all), effect))
#t1.all.order <- t1.all[order(t1.all$effect),]
t1.all[["sign"]] = ifelse(t1.all[["effect"]] >= 0, "secretor", "non-secretor")
par(mfrow=c(1,1))
t1.effect.plot <- ggplot(t1.all, aes(x=spp, y=effect, fill=sign))+geom_bar(stat="identity", width=0.5)+coord_flip()+theme_classic()+scale_fill_discrete(name="Enriched")
t1.effect.plot
```

### Comparison of secretors and non-secretors at Timepoint 2
The timepoint 2 pregnancies show a similar profile with respect to an increase in the amount of L.iners in the non-secretors.

```{r}
par(mfrow=c(1,2))
aldex.plot(t2.all, type="MA", test="welch")
aldex.plot(t2.all, type="MW", test="welch")
t2.all$spp <- with(t2.all, reorder(rownames(t2.all), effect))
t2.all[["sign"]] = ifelse(t2.all[["effect"]] >= 0, "secretor", "non-secretor")
par(mfrow=c(1,1))
t2.effect.plot <- ggplot(t2.all, aes(x=spp, y=effect, fill=sign))+geom_bar(stat="identity", width=0.5)+coord_flip()+theme_classic()+scale_fill_discrete(name="Enriched")
t2.effect.plot
```

### Comparison of Lactobacillus depleted secretors and non-secretors at timepoint 1

```{r}
t1asvDEP <- read.csv("T1CountsDAALDep.csv", header=T, row.names=1)
t1condsDEP <- c(rep("NSec",15), rep("Sec",47))
t2asvDEP <- read.csv("T2CountsDAALDep.csv", header=T, row.names=1)
t2condsDEP <- c(rep("NSec",13), rep("Sec",40))
t1DEP.aldex <- aldex.clr(t1asvDEP, t1condsDEP, mc.samples=1000, verbose=T)
t1DEP.aldex.tt <- aldex.ttest(t1DEP.aldex, paired.test=F, verbose=T)
t1DEP.aldex.effect <- aldex.effect(t1DEP.aldex, verbose=F)
t1DEP.all <- data.frame(t1DEP.aldex.tt, t1DEP.aldex.effect, stringsAsFactors = F)
t2DEP.aldex <- aldex.clr(t2asvDEP, t2condsDEP, mc.samples=1000, verbose=T)
t2DEP.aldex.tt <- aldex.ttest(t2DEP.aldex, paired.test=F, verbose=T)
t2DEP.aldex.effect <- aldex.effect(t2DEP.aldex, verbose=F)
t2DEP.all <- data.frame(t2DEP.aldex.tt, t2DEP.aldex.effect, stringsAsFactors = F)
```

###Plots at timepoint 1

```{r}
par(mfrow=c(1,2))
aldex.plot(t1DEP.all, type="MA", test="welch")
aldex.plot(t1DEP.all, type="MW", test="welch")
t1DEP.all$spp <- with(t1DEP.all, reorder(rownames(t1DEP.all), effect))
t1DEP.all[["sign"]] = ifelse(t1DEP.all[["effect"]] >= 0, "secretor", "non-secretor")
par(mfrow=c(1,1))
t1DEP.effect.plot <- ggplot(t1DEP.all, aes(x=spp, y=effect, fill=sign))+geom_bar(stat="identity", width=0.5)+coord_flip()+theme_classic()+scale_fill_discrete(name="Enriched")
t1DEP.effect.plot
```

###Plots at timepoint 2

```{r}
par(mfrow=c(1,2))
aldex.plot(t2DEP.all, type="MA", test="welch")
aldex.plot(t2DEP.all, type="MW", test="welch")
t2DEP.all$spp <- with(t2DEP.all, reorder(rownames(t2DEP.all), effect))
t2DEP.all[["sign"]] = ifelse(t2DEP.all[["effect"]] >= 0, "secretor", "non-secretor")
par(mfrow=c(1,1))
t2DEP.effect.plot <- ggplot(t2DEP.all, aes(x=spp, y=effect, fill=sign))+geom_bar(stat="identity", width=0.5)+coord_flip()+theme_classic()+scale_fill_discrete(name="Enriched")
t2DEP.effect.plot
```

