---
title: "FUT2modelling"
author: "S.Kundu"
date: "02/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Parallel cross-sectional modelling

Modelling (GLMM) of the secretor data/vaginal microbiome data across three timepoints in pregnancy.

### 1.Modelling with Lactobacillus status at Timepoint 1 ("early" pregnancy)

Perform standard dummy coding regression using the GlmmTMB library to fit a random intercept model (ethnicity coded as a four level categorical variable [CSAsia, European, Afro-Caribbean and Other]) regressing gestational age (days) on the following covariates with a gamma distribution:

Age
BMI (fbmi):
  0 - 18.5 to 24.99
  1 - <18.5
  2 - 25.0 to 29.99
  3 - >30.0
Previous PTB/MTL (fpbd):
  0 - no previous PTB or MTL
  1 - Previous PTB or MTL
Cerclage (fstitch):
  0 - no cerclage
  1 - cerclage
Previous Lletz treatment(floop):
  0 - no previous Lletz
  1 - previous Lletz treatment
Ethnicity (feth):
  1 - Central/South Asian
  3 - European
  5 - Afro-Caribbean
  6 - East Asian/MENA/Mixed/Other/Unknown
Secretor status (fsec):
  0 - secretor
  1 - non-secretor
Lactobacillus status (ft1ld, ft2ld and ft3ld in early, mid and late pregnancy):
  0 - Lactobacillus dominated
  1 - Lactobacillus depleted
CST (ft1cst, ft2cst and ft3cst in early, mid and late pregnancy):
  1 - CST 1
  2 - CST 2
  3 - CST 3
  4 - CST 4
  5 - CST 5
ABO group:
  0 - A
  1 - AB
  2 - B
  3 - O

We apply a reflection transformation to the gestational age variable to use a gamma distribution ("lgest" variable).

Note
1. Lactobacillus status is defined as +/-90% Lactobacillus abundance
2. Our previous analyses (not shown here) indicate only moderate evidence of multicollinearity (VIF values are all low, <10) between some variables (Previous PTB/MTL and previous Lletz).

```{r}
load("mod.rdata")
library(glmmTMB)
library(car)
library(emmeans)
library(DHARMa)
#Run GLMM incoroporating Lactobacillus status at Timepoint 1 and secretor status
glmm_t1 <- glmmTMB(lGest~Age+fbmi+fpbd+fstitch+floop+fsec+ft1ld+fsec:ft1ld+(1|feth), family=Gamma(link="log"), data=mod)
#Perform analysis of deviance using the Anova (using Anova from the car package rather than the core anova function as the former automatically takes care of unbalanced designs)
Anova(glmm_t1)
summary(glmm_t1)
#Perform a simpler GLM (for comparison with the GLMM)
glm_t1 = glmmTMB(lGest~Age+fbmi+fpbd+fstitch+floop+fsec+ft1ld+fsec:ft1ld, family=Gamma(link="log"), data=mod)
AIC(glm_t1,glmm_t1)
simglmmT1 <- simulateResiduals(fittedModel=glmm_t1, plot=T)
testDispersion(simglmmT1)
emm1 <- emmeans(glmm_t1, specs = ~fsec:ft1ld, type="response")
emm1
contrast(emm1, method="pairwise", adjust = "none")
```

The analysis of deviance results indicate that Previous PTB/MTL, Cerclage, Previous Lletz and the secretor status:Lactobacillus interaction terms are significant covariates. Residual analysis using the Dharma package indicates that the model fits well for each covariate. The coefficients indicate that the non-secretors with a Lactobacillus depleted microbiome have shorter gestational lengths (hence positive relationship with the inverted variable) compared with the baseline (women who are secretors with Lactobacillus dominated microbiomes). Women who have had a previous PTB/MTL also have shorter gestational lengths (compared with women who have not) as do women who get a cervical stitch (cerclage). Comparing the random intercept model with the GLM indicates that the latter is actually a slightly better fit (lower AIC), i.e., ethnicity does not appear to be a major factor (bias towards European women in this cohort may explain lack of signal with respect to ethnicity).

Post hoc comparison using the emmeans package suggests that the most significant comparison is between the non-secretors with Lactobacillus depleted microbiomes and non-secretors with Lactobacillus dominant microbiomes (comparison of secretors that have Lactobacillus depleted microbiomes with non-secretors that also have Lactobacillus deplete microbiomes is close to significance).

### 2.Modelling with Lactobacillus status at Timepoint 2 ("mid" pregnancy)

```{r}
glmm_t2 <- glmmTMB(lGest~Age+fbmi+fpbd+fstitch+floop+fsec+ft2ld+fsec:ft2ld+(1|feth), family=Gamma(link="log"), data=mod)
summary(glmm_t2)
Anova(glmm_t2)
glm_t2 = glmmTMB(lGest~Age+fbmi+fpbd+fstitch+floop+fsec+ft2ld+fsec:ft2ld, family=Gamma(link="log"), data=mod)
AIC(glm_t2,glmm_t2)
simglmmT2 <- simulateResiduals(fittedModel=glmm_t2, plot=T)
testDispersion(simglmmT2)
emm2 <- emmeans(glmm_t2, specs = ~fsec:ft2ld, type="response")
emm2
contrast(emm2, method="pairwise", adjust = "none")

```

The analysis of deviance indicates that at Timepoint 2 in oregnancy the interaction term is no longer significant - the conditional model shows that the coefficient is also less positive indicating a diminished reducing effect on gestational length.

### 3.Modelling timepoint 3 (late pregnancy)

```{r}
glmm_t3 <- glmmTMB(lGest~Age+fbmi+fpbd+fstitch+floop+fsec+ft3ld+fsec:ft3ld+(1|feth), family=Gamma(link="log"), data=mod)
summary(glmm_t3)
Anova(glmm_t3)
glm_t3 = glmmTMB(lGest~Age+fbmi+fpbd+fstitch+floop+fsec+ft3ld+fsec:ft3ld, family=Gamma(link="log"), data=mod)
AIC(glm_t3,glmm_t3)
simglmmT3 <- simulateResiduals(fittedModel=glmm_t3, plot=T)
testDispersion(simglmmT3)
emm3 <- emmeans(glmm_t3, specs = ~fsec:ft3ld, type="response")
emm3
contrast(emm3, method="pairwise", adjust = "none")

```

The analysis of deviance for the GLMM at Timepoint 3 indicates that the interaction term is not significant and the coefficient is now negative (indicating an increasing, rather than decreasing, effect on gestational length). At this point many of the preterm births have delivered, i.e., non-random missingness.

### 4.Modelling with CST at Timepoint 1 ("early" pregnancy)

Repeat the first model, but this time with the CST rather than Lactobacillus status. Here L.crispatus dominated microbiome (CST 1) will act as the baseline.

```{r}
glmm_cst <- glmmTMB(lGest~Age+fbmi+fpbd+fstitch+floop+fsec+ft1cst+fsec:ft1cst+(1|feth), family=Gamma(link="log"), data=mod)
Anova(glmm_cst)
summary(glmm_cst)
simglmmcst <- simulateResiduals(fittedModel=glmm_cst, plot=T)
testDispersion(simglmmcst)
emmcst <- emmeans(glmm_cst, specs = ~fsec:ft1cst, type="response")
emmcst
contrast(emmcst, method="pairwise", adjust="none")
```

The Analysis of deviance indicates that neither CST status (in early pregnancy) nor secretor status are, by themselves, significant explanatory variables of gestational length. However, the interaction term between CST and secretor status is statistically significant.

The conditional model highlights non-secretors with CSTs 2, 3 and 4 as having shorter gestational lengths compared to gestation of the baseline patient (intercept): secretor with CST 1 microbiome in early pregnancy (and typical BMI, no stitch, no Loop excision and no previous PTB/MTL). Additionally, the conditional model highlights the non-secretors with  CST 1 microbiota (in early pregnancy) as having significantly longer gestational lengths compared to the CST 1 secretors (baseline).

##ABO status (at timepoint 1)

###Modelling ABO status on gestational length (including secretor status)

Repeat the previous model but include secretor status as an additional fixed effect.

```{r}
glmm_t1_abo <- glmmTMB(lGest~Age+fbmi+fpbd+fstitch+floop+fsec+fabo+ft1ld+fsec:ft1ld+(1|feth), family=Gamma(link="log"), data=mod)
Anova(glmm_t1_abo)
summary(glmm_t1_abo)
glmm_t2_abo <- glmmTMB(lGest~Age+fbmi+fpbd+fstitch+floop+fsec+fabo+ft2ld+fsec:ft2ld+(1|feth), family=Gamma(link="log"), data=mod)
Anova(glmm_t2_abo)
summary(glmm_t2_abo)
glmm_t3_abo <- glmmTMB(lGest~Age+fbmi+fpbd+fstitch+floop+fsec+fabo+ft3ld+fsec:ft3ld+(1|feth), family=Gamma(link="log"), data=mod)
Anova(glmm_t3_abo)
summary(glmm_t3_abo)
```

The ANOVA of the model indicates that ABO is not a significant explanatory variable of gestational length. The Secretor status:Lactobacillus interaction term is still significant.
