---
title: "RFanalyses"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Random Forest analyses of FUT2 associated microbiome data
Comparison of the microbial composition between:
1. Secretors and non-secretors (early and mid-pregnancy)
2. LD- secretors and LD- non-secretors (early pregnancy)
3. Non-secretor PTB and term (early pregnancy)

### Steps required
1. Generate a mapping file for each of the above comparisons:
a) Samples Id (PREM) mapped to secretor status (early)
b) As above but mid
c) Samples ID (PREM) mapped to secretor status for LD- samples in early pregnancy
d) Samples ID (PREM) mapped to term and PTB (for just the non-secretors)
e) Generate a CLR transformed count table for the entire dataset
e) Use the sample list in each of the above tables to generate CLR-transformed ASV count tables for each of the comparisons (use Python Pandas to create the CLR transformed count files for each analysis).

```{r}
library(zCompositions)
allsp <- read.csv("allspcounts.csv", header=T, row.names = 1)
allsp.czm <- cmultRepl(t(allsp), label=0, method = "CZM")
allsp.clr <- t(apply(allsp.czm, 1, function(x){log(x)-mean(log(x))}))
fin.allsp.clr <- t(allsp.clr)
write.csv(fin.allsp.clr, file="AllSpCLRCounts.csv")
```

###RF comparison of microbiome in LD- non-secretors versus LD- secretors
Note: set ntree paramter to something higher, e.g., 10001, for a more robust trees. Also include a permutation test to test for significance.
TODO: plot the barplot in GGplot rather than in base graphics.

```{r}
library(plyr)#for the arrange function
library(randomForest)#for the random forest classification and regression
library(rfUtilities)#for the permutation test of significance
t1dep_meta <- read.csv("T1CLRLDepSecSt_meta.csv", header=T, row.names = 1)
t1dep_meta$fSec <- as.factor(t1dep_meta$SecretorStatus)
t1dep_asv <- read.csv("T1CLRLDepSecSt_counts.csv", header=T, row.names = 1)
colnames(t1dep_asv)==rownames(t1dep_meta)
asv_table_scaled_state <- data.frame(t(t1dep_asv))
asv_table_scaled_state$state <- t1dep_meta[rownames(asv_table_scaled_state), "fSec"]
set.seed(151)
RF_state_classify <- randomForest(x=asv_table_scaled_state[,1:(ncol(asv_table_scaled_state)-1)] , y=asv_table_scaled_state[ , ncol(asv_table_scaled_state)] , ntree=10001, importance=TRUE, proximities=TRUE)
RF_state_classify_sig <- rf.significance( x=RF_state_classify ,  xdata=asv_table_scaled_state[,1:(ncol(asv_table_scaled_state)-1)] , nperm=1000 , ntree=10001)#check for significance
RF_state_classify
RF_state_classify_imp <- as.data.frame( RF_state_classify$importance )
RF_state_classify_imp$features <- rownames( RF_state_classify_imp )
RF_state_classify_imp_sorted <- arrange( RF_state_classify_imp  , desc(MeanDecreaseAccuracy))
barplot(RF_state_classify_imp_sorted$MeanDecreaseAccuracy, ylab="Mean Decrease in Accuracy (Variable Importance)", main="RF Classification Variable Importance Distribution")
barplot(RF_state_classify_imp_sorted[1:10,"MeanDecreaseAccuracy"], names.arg=RF_state_classify_imp_sorted[1:10,"features"] , ylab="Mean Decrease in Accuracy (Variable Importance)", las=2, ylim=c(0,0.02), main="Classification RF")
```

###RF comparison of microbiome composition between term and PTB non-secretors in Early pregnancy
As above, set ntree to 10001 for more robust trees

```{r}
t1gest_asv <- read.csv("T1NSecGest_counts.csv", header=T, row.names = 1)
t1gest_meta <- read.csv("T1NSecGest_meta.csv", header=T, row.names = 1)
t1gest_meta$fOut <- as.factor(t1gest_meta$Outcome)
colnames(t1gest_asv)==rownames(t1gest_meta)
asv_table_out <- data.frame(t(t1gest_asv))
asv_table_out$out <- t1gest_meta[rownames(asv_table_out), "fOut"]
set.seed(171)
RF_out_classify <- randomForest(x=asv_table_out[,1:(ncol(asv_table_out)-1)] , y=asv_table_out[ , ncol(asv_table_out)] , ntree=10001, importance=TRUE, proximities=TRUE)
RF_out_classify_sig <- rf.significance( x=RF_state_classify ,  xdata=asv_table_scaled_state[,1:(ncol(asv_table_scaled_state)-1)] , nperm=1000 , ntree=10001)
RF_out_classify
RF_out_classify_imp <- as.data.frame( RF_out_classify$importance )
RF_out_classify_imp$features <- rownames( RF_out_classify_imp )
RF_out_classify_imp_sorted <- arrange( RF_out_classify_imp  , desc(MeanDecreaseAccuracy))
barplot(RF_out_classify_imp_sorted$MeanDecreaseAccuracy, ylab="Mean Decrease in Accuracy (Variable Importance)", main="RF Classification Variable Importance Distribution")
barplot(RF_out_classify_imp_sorted[1:10,"MeanDecreaseAccuracy"], names.arg=RF_out_classify_imp_sorted[1:10,"features"] , ylab="Mean Decrease in Accuracy (Variable Importance)", las=2, ylim=c(0,0.02), main="Classification RF")
```

The above RF indicates little signal for separating the non-secretors that delivered term from those that deliver perterm (+/- 37 weeks).

###RF regression of gestational age on ASVs

```{r}
asv_table_gest <- data.frame(t(t1gest_asv))
asv_table_gest$gest <- t1gest_meta[rownames(asv_table_gest), "Gestational.age"]
set.seed(161)
RF_gest_regress <- randomForest( x=asv_table_gest[,1:(ncol(asv_table_gest)-1)] , y=asv_table_gest[ , ncol(asv_table_gest)] , ntree=10001, importance=TRUE, proximities=TRUE )
RF_gest_regress
RF_gest_regress_sig <- rf.significance( x=RF_state_classify ,  xdata=asv_table_scaled_state[,1:(ncol(asv_table_scaled_state)-1)] , nperm=1000 , ntree=10001 )
RF_gest_regress_imp <- as.data.frame( RF_gest_regress$importance )
RF_gest_regress_imp$features <- rownames( RF_gest_regress_imp )
RF_gest_regress_imp_sorted <- arrange( RF_gest_regress_imp  , desc(`%IncMSE`)  )
barplot(RF_gest_regress_imp_sorted$`%IncMSE`, ylab="% Increase in Mean Squared Error (Variable Importance)", main="RF Regression Variable Importance Distribution")
barplot(RF_gest_regress_imp_sorted[1:10,"%IncMSE"], names.arg=RF_gest_regress_imp_sorted[1:10,"features"] , ylab="% Increase in Mean Squared Error (Variable Importance)", las=2, ylim=c(0,0.7), main="Regression RF")
```

###RF comparison of microbiome between secretors and non-secretors
####Timepoint 1 (early pregnancy)

```{r}
setwd("~/Dropbox/FUT2/Analysis/New analyses/FinalCountFiles/RFAnalyses/NsecVSSec")
t1_allsp_asv <- read.csv("T1CLRCounts.csv", header=T, row.names = 1)
t1_allsp_meta <- read.csv("T1RFSecSt_meta.csv", header=T, row.names = 1)
t1_allsp_meta$fSec <- as.factor(t1_allsp_meta$SecretorStatus)
colnames(t1_allsp_asv)==rownames(t1_allsp_meta)
t1_allsp_table_sec <- data.frame(t(t1_allsp_asv))
t1_allsp_table_sec$Sec <- t1_allsp_meta[rownames(t1_allsp_table_sec), "Sec"]
set.seed(151)
RF_secT1_classify <- randomForest( x=t1_allsp_table_sec[,1:(ncol(t1_allsp_table_sec)-1)] , y=t1_allsp_table_sec[ , ncol(t1_allsp_table_sec)] , ntree=10001, importance=TRUE, proximities=TRUE)
RF_secT1_classify_sig <- rf.significance(x=RF_secT1_classify,  xdata=t1_allsp_table_sec[,1:(ncol(t1_allsp_table_sec)-1)] , nperm=1000 , ntree=10001)#check for significance
RF_secT1_classify
RF_secT1_classify_imp <- as.data.frame(RF_secT1_classify$importance)
RF_secT1_classify_imp$features <- rownames( RF_secT1_classify_imp )
RF_secT1_classify_imp_sorted <- arrange( RF_secT1_classify_imp  , desc(`%IncMSE`))
barplot(RF_secT1_classify_imp_sorted$`%IncMSE`, ylab="% Increase in Mean Squared Error (Variable Importance)", main="RF Regression Variable Importance Distribution")
barplot(RF_secT1_classify_imp_sorted[1:10,"MeanDecreaseAccuracy"], names.arg=RF_secT1_classify_imp_sorted[1:10,"features"] , ylab="Mean Decrease in Accuracy (Variable Importance)", las=2, ylim=c(0,0.02), main="Classification RF")
```

####Timepoint 2 (mid-pregnancy)

```{r}
t2_allsp_asv <- read.csv("T2CLRCounts.csv", header=T, row.names = 1)
t2_allsp_meta <- read.csv("T2RFSecSt_meta.csv", header=T, row.names = 1)
t2_allsp_meta$fSec <- as.factor(t2_allsp_meta$SecretorStatus)
colnames(t2_allsp_asv)==rownames(t2_allsp_meta)
t2_allsp_table_sec <- data.frame(t(t2_allsp_asv))
t2_allsp_table_sec$Sec <- t2_allsp_meta[rownames(t2_allsp_table_sec), "Sec"]
set.seed(151)
RF_secT2_classify <- randomForest(x=t2_allsp_table_sec[,1:(ncol(t2_allsp_table_sec)-1)] , y=t2_allsp_table_sec[ , ncol(t2_allsp_table_sec)] , ntree=10001, importance=TRUE, proximities=TRUE)
RF_secT2_classify_sig <- rf.significance(x=RF_secT2_classify ,  xdata=t2_allsp_table_sec[,1:(ncol(t2_allsp_table_sec)-1)] , nperm=1000 , ntree=10001 )#check for significance
RF_secT2_classify
RF_secT2_classify_imp <- as.data.frame( RF_secT2_classify$importance )
RF_secT2_classify_imp$features <- rownames(RF_secT2_classify_imp )
RF_secT2_classify_imp_sorted <- arrange(RF_secT2_classify_imp  , desc(`%IncMSE`))
barplot(RF_secT2_classify_imp_sorted$`%IncMSE`, ylab="% Increase in Mean Squared Error (Variable Importance)", main="RF Regression Variable Importance Distribution")
barplot(RF_secT2_classify_imp_sorted[1:10,"MeanDecreaseAccuracy"], names.arg=RF_secT2_classify_imp_sorted[1:10,"features"] , ylab="Mean Decrease in Accuracy (Variable Importance)", las=2, ylim=c(0,0.02), main="Classification RF")
```

